*Author : Chau Nam Ky Nguyen*  
*Date : 14th November, 2014*  
*Course: Reproduciple research* 

---------------------------------------------------------

### Loading and preprocessing the data

1.Loading data

```{r}
data <- read.csv ("D:/Work/DataScience/05_Reproducible_Research/repdata-data-activity/activity.csv")
summary (data)
```
2.Preprocessing the data: Convert the type of the column "date" from "Factor" to "Date"
```{r}
date_tmp <- as.Date(data$date)
data$date <- date_tmp 
str (data)
```

### What is mean total number of steps taken per day?

1.Make a histogram of the total number of steps taken each day

```{r}
sum_result <- aggregate (steps ~ date, data = data, FUN = sum)
par (mfrow = c(1,1))
hist (sum_result$steps, xlab = "Steps", ylab = "Number of Date", col = "Blue", main = "")
```

2.Calculate and report the mean and median total number of steps taken per day

```{r}
median_result <- aggregate (steps ~ date, data = data, FUN = median)
mean_result   <- aggregate (steps ~ date, data = data, FUN = mean)
total_result  <- merge (mean_result, median_result, by = c ("date","date"))
names (total_result) <- c("date","mean","median")
head (total_result)
```


### What is the average daily activity pattern?

1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
mean_interval <- aggregate (steps ~ interval, data = data, FUN = mean)
plot (mean_interval$step ~ mean_interval$interval,type = "l", xlab = "Interval", ylab = "Average Step", col = "Green")
```


2.Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
mean_interval[mean_interval$step == max (mean_interval$step),]
```

### Imputing missing values

1.Calculate and report the total number of missing values in the dataset

```{r}
a <- sum(is.na (data))
```

**The total of NA rows is `r a`**  

2.Devise a strategy for filling in all of the missing values in the dataset

*The Stratergy is : Imputing missing values by assigning the mean of all intervals to a missing date*

```{r}
na_date <- unique(data[is.na (data),2])

new_data <- data

for (i in na_date) 
    new_data [new_data$date == i,"steps"] = as.integer (mean_interval[,2])

new_sum_result <- aggregate (steps ~ date, data = new_data, FUN = sum)

```
3.Create a new dataset that is equal to the original dataset but with the missing data filled in.

*Show only previous missing date with new data*

```{r}
head (subset (new_data, date %in% na_date),5)
```


4a.Report the mean and median total number of steps taken per day

Calculate the new median and mean 

```{r}
new_median_result <- aggregate (steps ~ date, data = new_data, FUN = median)
new_mean_result   <- aggregate (steps ~ date, data = new_data, FUN = mean)
```

Add median and mean of missing days

```{r}
tmp <- as.data.frame (cbind(as.character (na_date), matrix (NA, 8,2)))

tmp[,1] <- as.Date    (tmp[,1])
tmp[,2] <- as.numeric (tmp[,2])
tmp[,3] <- as.numeric (tmp[,3])

names (tmp)          <- c("date","mean","medium")
names (total_result) <- c("date","mean","medium")

total_result <- rbind (total_result,tmp)
```

Report the mean and median of original and new data

```{r}

total_result <- total_result[order (total_result$date),]

total_result <- merge (total_result, new_mean_result  , by = c("date","date"))
total_result <- merge (total_result, new_median_result, by = c("date","date"))

names (total_result) <- c("date","old mean","old medium","new mean","new medium")

head (total_result)
```

4b. Make a histogram of the total number of steps taken each day and Calculate  

```{r}
par (mfrow = c(1,2))

hist (sum_result$steps    , xlab = "Steps", ylab = "Number of Date", col = "Blue", main = "Original")
hist (new_sum_result$steps, xlab = "Steps", ylab = "Number of Date", col = "Cyan", main = "New")
```

Do these values differ from the estimates from the first part of the assignment? 
*Yes, the date is different. Missing dates now are shown in the histogram*

What is the impact of imputing missing data on the estimates of the total daily number of steps?
*As the above figure, the number of days is increased from 25 to 35*


### Are there differences in activity patterns between weekdays and weekends?

1.Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day

Find weekday in data

```{r}
tmp <- weekdays (new_data$date)
tmp <- sapply (tmp,function(x) if (x %in% c("Saturday","Sunday")) "Yes" else "No")

new_data <- cbind (new_data, weekend = tmp)

weekend_data <- subset (new_data, weekend == "Yes")
weekday_data <- subset (new_data, weekend == "No")

```

Calculate the average of steps according to "Weekend" or "Weekday"

```{r}
weekday_mean <- aggregate (steps ~ interval, data = weekday_data, FUN = mean)
weekend_mean <- aggregate (steps ~ interval, data = weekend_data, FUN = mean)
```

Create a new variable that has 2 levels : "Weekend" and "Weekday" and plot the result

```{r}
weekend_mean <- cbind (weekend_mean, c ("Weekend")) 
weekday_mean <- cbind (weekday_mean, c ("Weekday"))

names (weekday_mean) <- c("interval","mean","date")
names (weekend_mean) <- c("interval","mean","date")

tmp <- rbind (weekday_mean, weekend_mean)

str (tmp)
```

2.Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r}
library (lattice)
px <- xyplot (mean ~ interval | date, data = tmp, type = "l", ylab = "Number of Steps", xlab = "Interval", layout = c(1,2))
print (px)
```
